Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex
abstract

%%%%%%%%% ABSTRACT
\begin{abstract}
    Using 3D model to detect and register the model to RGB image has been a growing field of study. Recently, Aubry \etal \cite{Aubry14} and J. Lim \etal \cite{Lim14} tried to estimate pose and align exact CAD model to an image using part-based templates. Malisiewicz \etal \cite{Malisiewicz11} also addresses the issue by transferring metadata after detection. Hejrati \etal \cite{Hejrati14} uses template synthesis to detect an object. All these approaches require extensive training, a lot of memory space. Also, the training based approach requires defining fixed viewpoints which limit themselves.
  To overcome these difficulties, we propose an efficient and fast way to synthesize and validate templates using on the fly realistic rendering that can jointly optimize scale, translation, rotation and focal length continuously. We render an object and extract features and decorrelate spatial dependencies within them on the fly to make a discriminative template. To speed up the decorrelation procedure, we adopt a gradient based algorithm which is numerically more stable than decomposition procedure. Finally, we efficiently search for candidate match by computing convolution using FFT and jointly optimize to match the object accurately. To speed up further, we propose an efficient GPU implementation and tested on PASCAL images.

  In contrast to other mid-level patch based 3D matching methods which requires extensive training and more than 10Gb of memory space and physical storage, ours require no training nor saving the templates thus enabling it to run on any personal computer. 
\end{abstract}

@@

\section{Related Works}
\subsection{Whitened Histograms of Orientations (WHO)}
% History of Whitening and how whitening works
Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO) \cite{Hariharan12}, which used statistics from natural images to whiten HOG templates, making them more discriminative. Whitening is a common signal processing operation for decorrelating a set of random variables \cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a $k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we remove 2nd order correlation between all features. So, all elements in the random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots, \tilde{X}_k\right]$, become more independent. In the context of HOG features, a whitened HOG template will produce Gaussian noise when convolved with a random natural image, but a delta-like response on an image matching the template. WHO templates are more discriminative than standard HOG templates because they only capture the gradients that differentiate the object from random images.

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)?}

% Using HOG
% In this paper, we also used Histograms of Orientations feature \cite{Dalal05} but in general, we could use any dense features.

@@

Our second innovation is 'Non-Zero' whitening. When synthesizing detection templates from rendered images, a common problem is how to handle the background. Some methods \scream{which?} render the model over a natural image background, but this results in gradients in the background being incorporated into the discriminative template. Alternatively, we could leave the background textureless \scream{did Hariharan do this?} (see Fig.~\ref{fig:rendering}), but whitening the resulting HOG template produces in undesirable background artifacts. These are created when centering the template (by subtracting the mean $\mu$), where strong negative weights are introduced in the textureless region (as seen in Fig.~\ref{fig:whocomparison}). This could result in positive matches being suppressed due to spurious background gradients.

@@

Therefore, assuming WSS allows us to synthesize the covariance matrix for templates \emph{of any size} from the autocovariance matrix. This hugely reduces the memory required because to build the covariance matrix of a template of $w \times h$ HOG cells (with 31 elements per cell), we only need a $w \times h \times 31$ autocovariance matrix. In comparison, the covariance matrix itself is $(w \times h \times 31) \times (w \times h \times 31)$. Furthermore, by using the autocovariance we can avoid storing the covariance matrix for every different aspect ratio used during detection.

To further simplify $\Gamma$, we assumed horizontal and vertical symmetry. In our implementation, we compute a $40 \times 40 \times 31$ autocovariance matrix, meaning we can synthesize covariance matrices for HOG grids as big as $40 \times 40$. In practice we limit the number of cells per template to 250.

We use the GPU to parallelize the lookup of the correct autocovariance cells during covariance matrix syntheses, which is \scream{x} times faster than using the CPU (Figure \ref{fig:covariancetime}).

@@

To whiten a signal $x$, we have to solve $\Sigma w = x - \mu$ which takes $O(n^3)$ time. In \cite{Hariharan12}, the authors make use of the fact that covariance matrices are Hermitian and positive semidefinite to solve the system via the Cholesky decomposition, which is twice as fast as the LU decomposition. However, by considering the properties of the covariance matrix of the HOG template, we show that the conjugate gradient method can further speed up the whitening stage.

%  If we use a high resolution template such as \ref{fig:rendering}

@@

%In our setting, we used high resolution template which has more than 250 HOG cells. This result in $\Sigma \in \mathcal{R}^{7500 \times 7500}$. It takes about 9 seconds if we use LU decomposition and 5 seconds if we use Cholesky decomposition.  % The default way of solving the linear equation in many linear algebra library is decomposing the matrix using LU Decomposition (in MATLAB the default method of backslash operator) which takes $O(\frac{2}{3}n^3)$ flops for $n\times n$ matrix. Since the matrix, $\Sigma + \lambda I$ is Hermitian and Positive Semi Definite. We can speed up using Cholesky Decomposition.

Figure \ref{fig:spectral} shows the singular values of a template covariance matrix with a small regularization constant added to the diagonal. The singular values are clustered in small region and most of the energies are concentrated on the first \scream{x} singular values. Therefore, conjugate gradient will converge in $O(n^2\kappa)$ time, faster

We also used iterative Conjugate Gradient to make use of low rank property of the covariance \cite{Gharbi12}. After adding small regularization, the regularized covariance matrix has nice spectral property (Figure \ref{fig:spectral}). The singular values are clustered in small region and most of the energies are concentrated on the first (number) singular values. Thus, the conjugate gradients stage doesn't go upto the size of the matrix to converge.

@@

\begin{figure}[h]
    \begin{center}
    \setlength\tabcolsep{0pt}
\begin{tabular}{cc}
  \includegraphics[width=0.45\linewidth]{car_3dobject/7.png} &   
  \includegraphics[width=0.45\linewidth]{car_3dobject/8.png}\\ [-15pt]
%   (a) & (b) \\[0pt]
  \includegraphics[width=0.45\linewidth]{car_3dobject/11.png} &   
  \includegraphics[width=0.45\linewidth]{car_3dobject/12.png}\\ [-5pt]
%  (c) & (d) \\[0pt]
\end{tabular}
\end{center}
\caption{Detection results on 3D-Object dataset \cite{Savarese07}, each of four images have (top left) original image, (top right) top detection rendering overlaid, (bottom right) false positives, (bottom left) true positives. Each with confidence score and colored bounding box}
  \label{fig:3dobject}
\end{figure}

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
A large body of recent work on object detection has focused on exploiting 3D CAD model databases to improve detection performance. Many of these approaches work by aligning exact 3D models to images using templates generated from renderings of the 3D models at a finite set of discrete viewpoints. The training procedures for these approaches, however, are very expensive and require gigabytes of memory and storage, and the viewpoint discretization hampers pose estimation performance.

We propose an efficient method for synthesizing templates from 3D models that runs on the fly -- that is, it quickly produces detectors for an arbitrary viewpoint of a 3D model without expensive dataset-dependent training or template storage. Given a 3D model and an arbitrary continuous detection viewpoint, our method synthesizes a discriminative template by extracting features from a rendered view of the object and decorrelating spatial dependences among the features. Our decorrelation procedure relies on a gradient-based algorithm that is more numerically stable than standard decomposition-based procedures, and we efficiently search for candidate detections by computing FFT-based template convolutions. Due to the speed of our template synthesis procedure, we are able to perform joint optimization on continuous scale, translation, rotation, and focal length. We provide an efficient GPU implementation of our algorithm, and we validate its performance on 3DObject dataset and PASCAL 2012 dataset.

%     Using 3D model to detect and register the model to RGB image has been a growing field of study. Recently, Aubry \etal \cite{Aubry14} and J. Lim \etal \cite{Lim14} tried to estimate pose and align exact CAD model to an image using part-based templates. Malisiewicz \etal \cite{Malisiewicz11} also addresses the issue by transferring metadata after detection. Hejrati \etal \cite{Hejrati14} uses template synthesis to detect an object. All these approaches require extensive training, a lot of memory space. Also, the training based approach requires defining fixed viewpoints which limit themselves.
%  To overcome these difficulties, we propose an efficient and fast way to synthesize and validate templates using on the fly realistic rendering that can jointly optimize scale, translation, rotation and focal length continuously. We render an object and extract features and decorrelate spatial dependencies within them on the fly to make a discriminative template. To speed up the decorrelation procedure, we adopt a gradient based algorithm which is numerically more stable than decomposition procedure. Finally, we efficiently search for candidate match by computing convolution using FFT and jointly optimize to match the object accurately. To speed up further, we propose an efficient GPU implementation and tested on PASCAL images.

%   In contrast to other mid-level patch based 3D matching methods which requires extensive training and more than 10Gb of memory space and physical storage, ours require no training nor saving the templates thus enabling it to run on any personal computer. 
\end{abstract}

@@

\section{Related Works}
\subsection{Whitened Histograms of Orientations (WHO)}
% History of Whitening and how whitening works
Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO) \cite{Hariharan12}, which used statistics from natural images to whiten HOG templates, making them more discriminative. Whitening is a common signal processing operation for decorrelating a set of random variables \cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a $k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we remove 2nd order correlation between all features. So, all elements in the random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots, \tilde{X}_k\right]$, become more independent. In the context of HOG features, a whitened HOG template will produce Gaussian noise when convolved with a random natural image, but a delta-like response on an image matching the template. WHO templates are more discriminative than standard HOG templates because they only capture the gradients that differentiate the object from random images. \scream{chris added the following} Hariharan \etal. also interpreted the whitening as Linear Discriminant Analysis (LDA) where all classes have different means but have the same covariance. The whitening and LDA have the same formulation when we make a decision boundary $w = \Sigma^{-1}(\mu_+ - \mu_0)$ where $\mu_+$ is the features from a class (LDA) or a signal to whiten (whitening), $\mu_0$ is the negative class center (LDA) or a signal statistics (whitening) and $\Sigma$ is the covariance matrix of classes (LDA) or correlation within signal (whitening).

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)?}
% Using HOG
% In this paper, we also used Histograms of Orientations feature \cite{Dalal05} but in general, we could use any dense features.

@@

Our second innovation is 'Non-Zero' whitening. When synthesizing detection templates from rendered images, a common problem is how to handle the background. If we render the model over a natural image background, gradients in the background will be incorporated into the discriminative template. Alternatively, we could leave the background textureless \scream{did Hariharan do this?} (see Fig.~\ref{fig:rendering}), but whitening the resulting HOG template produces in undesirable background artifacts. These are created when centering the template (by subtracting the mean $\mu$), where strong negative weights are introduced in the textureless region (as seen in Fig.~\ref{fig:whocomparison}). This could result in positive matches being suppressed due to spurious background gradients.

@@

Therefore, assuming WSS allows us to synthesize the covariance matrix for templates \emph{of any size} from the autocovariance. This hugely reduces the memory required because to build the covariance matrix of a template of $w \times h$ HOG cells (with 31 elements each), we only need a $w \times h \times 31$ autocovariance matrix. In comparison, the covariance matrix itself is $(w \times h \times 31) \times (w \times h \times 31)$. Furthermore, by using the autocovariance we can avoid storing the covariance matrix for every different aspect ratio used during detection.

To further simplify $\Gamma$, we assumed horizontal and vertical symmetry. In our implementation, we compute a $40 \times 40 \times 31$ autocovariance matrix, meaning we can synthesize covariance matrices for HOG grids of as big as $40 \times 40$.

Synthesizing the covariance matrix is memory copy operations so we used GPU to speed up the process (Figure \ref{fig:covariancetime}).

@@

To whiten a signal $x$, we have to solve $\Sigma w = x - \mu$ which takes $O(n^3)$ time. In \cite{Hariharan12}, the authors make use of positive semidefinite and Hermitian properties of a covariance matrix to use Cholesky decomposition and Gaussian elimination which is twice as fast a the LU decomposition. However, if we make use of a spectral property of the covariance matrix, we can further speed up the whitening stage.

  If we use a high resolution template such as \ref{fig:rendering}

@@

In our setting, we used high resolution template which has more than 250 HOG cells. This result in $\Sigma \in \mathcal{R}^{7500 \times 7500}$. It takes about 9 seconds if we use LU decomposition and 5 seconds if we use Cholesky decomposition.  % The default way of solving the linear equation in many linear algebra library is decomposing the matrix using LU Decomposition (in MATLAB the default method of backslash operator) which takes $O(\frac{2}{3}n^3)$ flops for $n\times n$ matrix. Since the matrix, $\Sigma + \lambda I$ is Hermitian and Positive Semi Definite. We can speed up using Cholesky Decomposition.

We also used iterative Conjugate Gradient to make use of low rank property of the covariance \cite{Gharbi12}. After adding small regularization, the regularized covariance matrix has nice spectral property (Figure \ref{fig:spectral}. The singular values are clustered in small region and most of the energies are concentrated on the first (number) singular values. Thus, the conjugate gradients stage doesn't go upto the size of the matrix to converge.

@@

\begin{figure}[h]
    \begin{center}
    \setlength\tabcolsep{0pt}
\begin{tabular}{cc}
  \includegraphics[width=0.45\linewidth]{car_3dobject/7.png} &   
  \includegraphics[width=0.45\linewidth]{car_3dobject/8.png}\\ [-15pt]
%   (a) & (b) \\[0pt]
  \includegraphics[width=0.45\linewidth]{car_3dobject/11.png} &   
  \includegraphics[width=0.45\linewidth]{car_3dobject/12.png}\\ [-5pt]
%  (c) & (d) \\[0pt]
\end{tabular}
\end{center}
\caption{Detection results on 3D-Object dataset \cite{Savarese07}. Original image (left) and detection result overlaid on top with a bounding box and corresponding confidence score(right)}
  \label{fig:3dobject}
\end{figure}
